Toward a secure TCP/IP stack --- Technical report

### 2020-07-19 report/report.tex
Continue report

@@

@@

    \subsubsection{Concurrency: asynchronous mechanism}

@@

    Finally, all the result found by klee has been reported in the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
    This function is essential for all the verification of concurrent parts.

@@

    \subsubsection{User functions}

    Code written in SPARK to check if the transitions are coherent inside the user functions.

    \subsubsection{Concurrency}

    Modeling the concurrency and the interleaving semantic has been the hardest work.
    The held solution is to have "ghost" code that model the different transition that
    might have happened when the mutex is released.

    The interleaving semantic is also considered for synchronous events, when the code
    wait for a particular event.

    If we consider the state transition given by the code that receive the message, we
    are driven to consider a transition relation, denotes as $\rightarrow$. This relation
    is reflexive because a received message can be drop if it doesn't contain useful
    information. Then, for all TCP state $S$, $S \rightarrow S$. As multiples messages
    can be received, the transitive closure of $\rightarrow$ is considered:
    $\rightarrow^* = \bigcup_{n\in\mathbb{N}^*} \rightarrow^n$ which can be simplified to:
    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
    since every state is only reachable by a state at a distance at most of 3, which make
    the calculation of the transition closure easier by SPARK once we have $\rightarrow$ by
    unrolling loops.

>>>>

\usepackage[ruled]{algorithm2e}

@@

    \tableofcontents

@@

    \subsubsection{Overview of the concurrency challenge}

@@

    Finally, all the result found by klee has been reported in the function \lstinline[language=Ada]{Tcp_Process_One_Segment}.
    This function is essential for all the verification of concurrent parts and is reused everywhere
    the concurrency is involved to compute the results.

@@

    \subsubsection{Concurrency: asynchronous changes of state}

    Between the user functions calls, segment can be received and these receptions can drive to
    change of state of the socket. Between two function calls, an infinite number of segments
    can be potentially receive. The reception of one or zero segment is modeled by the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} in term of change of states. Then we have
    to consider the iteration of \lstinline[language=Ada]{Tcp_Process_One_Segment} in order to
    compute the result of the reception of multiple message.

    \begin{algorithm}[t]
        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Process\_Segment}}
        \SetKwProg{Fn}{function}{}{end}
        \Fn{\ProcessSegm{Socket}}{
            $S_{last} := \text{\textit{Socket}}$\;
            $S := S_{last}$\;
            \For{$i=1$ \KwTo $3$}{
                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
                $S := S \cup S_{last}$\;
            }
            \KwRet{$S$}\;
        }
        \caption{Reflexive and transitive closure of \lstinline[language=Ada]{Tcp_Process_One_Segment}}
        \label{algo:tcpProcessSegment}
    \end{algorithm}

    Let $\rightarrow$ be the transition function in the TCP automaton restrited to the transitions
    that can be performed by the reception of a message and its automatic response mechanism.
    Then we have $\rightarrow = \text{\lstinline[language=Ada]{Tcp_Process_One_Segment}}$ and
    we need to consider the reflexive transitive closure $\rightarrow^*$ of $\rightarrow$, with
    \[\rightarrow^* = \bigcup_{n\in\mathbb{N}} \rightarrow^n\]
    Now, by examining the TCP automaton in figure \ref{fig:TCPAutomaton}, we see that all state is
    only reachable by a state at a distance less than 3 (the maximum path is between SYN-SENT and
    CLOSE-WAIT if we pass by state SYN-RECEIVED and ESTABLISHED) without action by the user. Since
    we only consider the transitive closure in term of states, we can significantly reduce the number
    of iteration of $\rightarrow$ to compute $\rightarrow^*$ and finally we get:
    \[\rightarrow^* = \bigcup_{n=1}^3 \rightarrow^n\]
    It follows that algorithm \ref{algo:tcpProcessSegment} is enough good to compute the transitive closure of the function
    \lstinline[language=Ada]{Tcp_Process_One_Segment} by taking advantage of the fact that SPARK can unroll
    small loops to obtain more precise results.

    In the user functions, we have added a call to \lstinline[language=Ada]{Tcp_Process_Segment} everywhere
    there is a call to \lstinline{Os_Acquire_Mutex (Net_Mutex)}. Doing that helps to ensure that all the
    possible that all the possible input states are have been considered.

    \subsubsection{Concurrency: Synchronous exchange}

    The other mechanism to deal with concurrency is probably more difficult to apprehend and the
    use of the function \lstinline[language=Ada]{Tcp_Process_One_Segment} is even more noticeable.

    The C code contains a function \lstinline[language=C]{tcpWaitForEvents} in the file \texttt{tcp\_misc.c}
    that checks if the event is true when the function is called, by calling the function
    \lstinline[language=C]{tcpUpdateEvents}. If the wait is not completed at the time the function is
    called, then the mutex that was previously locked is released until the event become waited happen.
    Then, everything can happen meanwhile.

    In the code, the function \lstinline[language=C]{tcpUpdateEvents} is called at different locations
    and can raise the desired event. In particular this function is called each time the state of
    the socket is changed. This is sufficient for our purpose since we are only interested in the state
    changes.

    When a segment is received, the function \lstinline[language=C]{tcpUpdateEvents} is called if a
    change of state happens. Then, we can consider the algorithm \ref{algo:waitForEvents} to compute the set of possible
    final states after the wait. \textit{(This is done in the function \lstinline[language=Ada]{Tcp_Wait_For_Events_Proof},
    the function dedicated to proof in file \texttt{tcp\_misc\_binding.adb}.)}

    \begin{algorithm}[t]
        \SetKwFunction{ProcessSegm}{\lstinline[language=Ada]{Tcp\_Wait\_For\_Events}}
        \SetKwProg{Fn}{function}{}{end}
        \Fn{\ProcessSegm{Socket, Event, Event\_Mask}}{
            $S_{last} := \text{\textit{Socket}}$\;
            $S := S_{last}$\;
            $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
            \If{$E \wedge \text{Event\_Mask} \neq 0$}{
                \KwRet{$S$}\;
            }
            \For{$i=1$ \KwTo $3$}{
                $S_{last} :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$) \;
                $S := S \cup S_{last}$\;
                $E :=$ \lstinline[language=Ada]{Tcp_Process_One_Segment}($S_{last}$)\;
                \If{$E \wedge \text{Event\_Mask} \neq 0$}{
                    \KwRet{$S$}\;
                }
            }
            \KwRet{$\emptyset$}\;
        }
        \caption{Function to compute the possible state after when waiting for a particular event.}
        \label{algo:waitForEvents}
    \end{algorithm}

    We can compute precisely the state reach for each event waited thanks to the fact that SPARK
    unroll loops.

    \textcolor{red}{\paragraph{Weakness} If two threads want to lock the same mutex at the same time, which one win?}

    \section{Future work}
